---
title: "Amsterdam: A Data Scientist's City Guide"
author: "Sonja Sahebzad"
date: "September 16, 2026"
output:
  html_document:
    self_contained: true
    code_folding: hide
    theme: flatly
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(leaflet)
```

<style>
body {
  font-family: "Segoe UI", Calibri, Arial, sans-serif;
  color: #1f2937;
}
h1, h2 { color: #123b65; }
.subtitle {
  padding: 14px 18px;
  margin: 18px 0;
  background: #eef6fc;
  border-left: 5px solid #2171b5;
  border-radius: 4px;
}
.leaflet-container { border: 1px solid #b8c8d8; border-radius: 7px; }
</style>

<div class="subtitle">
This interactive map presents a small selection of places where learning, culture, and green space meet in Amsterdam. Select a marker to read about the location, switch the base map, or use the layer control to focus on one theme.
</div>

## Interactive map

```{r amsterdam-map, fig.width=11, out.width="100%"}
places <- data.frame(
  name = c(
    "Amsterdam Public Library (OBA)", "NEMO Science Museum",
    "University of Amsterdam", "Rijksmuseum", "Van Gogh Museum",
    "A'DAM Lookout", "Vondelpark", "ARTIS"
  ),
  category = c(
    "Learning & Science", "Learning & Science", "Learning & Science",
    "Culture", "Culture", "Culture", "Green Amsterdam", "Green Amsterdam"
  ),
  latitude = c(52.3758, 52.3741, 52.3680, 52.3600, 52.3584, 52.3838, 52.3579, 52.3660),
  longitude = c(4.9083, 4.9123, 4.8903, 4.8852, 4.8811, 4.9023, 4.8686, 4.9165),
  description = c(
    "A modern public library with study areas and broad city views.",
    "A hands-on science museum in Renzo Piano's waterfront building.",
    "A major research university in the historic centre of Amsterdam.",
    "The national museum of Dutch art and history.",
    "A museum dedicated to the work and life of Vincent van Gogh.",
    "A panoramic viewpoint across the IJ from Amsterdam Centraal.",
    "Amsterdam's best-known urban park and a calm place to recharge.",
    "The city's historic zoo, botanical setting, and science centre."
  ),
  stringsAsFactors = FALSE
)

group_colours <- c(
  "Learning & Science" = "#08519c",
  "Culture" = "#3182bd",
  "Green Amsterdam" = "#31a354"
)

city_map <- leaflet(options = leafletOptions(minZoom = 11)) |>
  addProviderTiles(providers$CartoDB.Positron, group = "Light map") |>
  addProviderTiles(providers$OpenStreetMap.Mapnik, group = "Street map") |>
  setView(lng = 4.8952, lat = 52.3702, zoom = 13)

for (theme_name in names(group_colours)) {
  selected <- places[places$category == theme_name, ]
  popup_text <- paste0(
    "<div style='font-family:Segoe UI,Arial;min-width:210px'>",
    "<strong style='color:#123b65;font-size:15px'>", selected$name, "</strong>",
    "<br><span style='color:", group_colours[[theme_name]], ";font-weight:600'>",
    selected$category, "</span><br><br>", selected$description, "</div>"
  )

  city_map <- city_map |>
    addCircleMarkers(
      data = selected,
      lng = ~longitude,
      lat = ~latitude,
      radius = 8,
      color = "white",
      weight = 2,
      fillColor = group_colours[[theme_name]],
      fillOpacity = 0.9,
      popup = popup_text,
      label = ~name,
      group = theme_name,
      labelOptions = labelOptions(
        direction = "auto",
        style = list("font-family" = "Segoe UI, Arial", "font-weight" = "600")
      )
    )
}

city_map |>
  addLegend(
    position = "bottomright",
    colors = unname(group_colours),
    labels = names(group_colours),
    title = "Place type",
    opacity = 0.9
  ) |>
  addLayersControl(
    baseGroups = c("Light map", "Street map"),
    overlayGroups = names(group_colours),
    options = layersControlOptions(collapsed = FALSE)
  )
```

## About this map

The map was created in R with the **leaflet** package. It uses open map tiles and a small curated dataset of Amsterdam locations. Hovering over a marker reveals its name; selecting it opens a short description. The layer control lets the reader compare themes without losing the geographic context.

